MLH1 Results Outline

### 2020-02-23 doc/Results.rmd
more calqs from Q2 sc length

<and the short biv>
Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).

@@

```{r Q1.glm_totalSC, echo=FALSE, message=FALSE}
#total-SC
#Q1.lmer_totSC.mouse.av.table

@@

```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE, message=FALSE}

@@

The following effects had significant p values given the above model

@@

From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).

@@

In the M2 model, the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).

<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-AE.  >

In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.

@@

```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}

Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
```

@@

#mouse av
Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length")+facet_wrap(~sex)+ylim(c(75,125) )

@@

```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE, eval=FALSE}

#REAL.long_mouse_table

#removed these plots for space
Q1_cell_long.biv <- ggplot(REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")

                        n_obs = length(Obj.ID),
                         mean.sc = mean(chromosomeLength)
                         )

cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)

#cell.avlong.biv <- ggplot( (REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")

#mouse av for long biv

mouse.show.long.biv <- ggplot((REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av. \nMale Long Biv SC Lengths")

mouse.show.long.biv

```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}

Q1.real.long_mouse_table <- REAL.long_mouse_table

###
lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)

Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value

Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
sex.results =  Q1lmer.MM_SC_long_sex, 
               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
Q1lmer.MM_SC_long_strain)
```

@@

motivation:
the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)

@@

@@

In order to test if SC lengths (SC area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.

total SC for males (which are), X additional cells were added to the male specific total SC data set described above

-reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice

@@

```{r Q2.SC.juvimice.list.copy, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

ppmm <- OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$mouse  %in% juvi_mice, ]   
```

@@

@@

musc and mol ttests
p = 
p=

@@

**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`). (the summary of mouse averages for this subset of ) **

>>>>

Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).

@@

```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
#total-SC
#Q1.lmer_totSC.mouse.av.table

@@

```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}

@@

The following effects had significant p values for the **total SC** from the above model

@@

Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).

@@

In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).

<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >

<In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>

<1. the longer female sc's is consistant and seen across all strains,
even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>

@@

@@

@@

@@

<motivation:
the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>

@@

while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR

@@

-reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice. (Since comparing within sex, there are no sex chromosomes differences and the long bivalent dataset can be used.

@@

@@

#mUSC-short
Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)

#mol-short
#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)

#musc-long
Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)

#mol-long
#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)

#not enought observations

```{r Q2.SC_HvL_allBiv, echo=FALSE}

#High-Low_All-BivData

Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)

#short
Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)

#long
Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)

@@

musc and mol ttests
mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means

musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`
the total sc mouse means are not significant for the high and low strains in musc

@@

**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`). (the summary of mouse averages for this subset of ) **

within subspecies

musc-short

musc-long

mol-short

mol-long

```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}

#Dom-Musc
Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
#p-value = 0.5

#Dom-Mol
Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
#p-value = 0.1
```

```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}

#DOM-musc-long
Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
#p-value = 0.08

#DOM-musc-short
Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
#p-value = 0.8

#DOM-mol-long
Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
#p-value = 0.4

#DOM-mol-short
Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
#p-value = 0.09

```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}

#DOM-MUSC-ALL
Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 4e-13

#DOM-MOL-ALL
Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 6e-10

#DOM-musc-long
Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.2

#DOM-musc-short
Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.8

#DOM-mol-long
Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.4

#DOM-mol-short
Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.06

across subspecies:

dom - musc

dom-mol
